# Matching Individual Items

While the generate pipeline is usually sufficient, it's sometimes important to tailor stimuli more precisely. For instance, it may be important that a matched word is a plausible replacement for a target word in a sentence. The `match_item()` function exists for this purpose.

## Example

Here's an example usage of `match_item()`, to suggest a word matched for "elephant" in terms of:

* Length (exactly)
* Frequency (within ±0.25 Zipf)
* Imageability (within ±1, on a 1-7 Likert rating scale)
* Part of Speech (i.e. is also a noun)

```{r, eval=FALSE}
library(LexOPS)

suggested_matches <- lexops %>%
  match_item(
    "elephant",
    Length,
    Zipf.SUBTLEX_UK = -0.25:0.25,
    IMAG.Glasgow_Norms = -1:1,
    PoS.SUBTLEX_UK
  )
```

```{r, echo=FALSE, message=FALSE}
library(LexOPS)

suggested_matches <- lexops %>%
  match_item(
    "elephant",
    Length,
    Zipf.SUBTLEX_UK = -0.25:0.25,
    IMAG.Glasgow_Norms = -1:1,
    PoS.SUBTLEX_UK
  ) %>%
  dplyr::select(string, euclidean_distance, Length, Zipf.SUBTLEX_UK, IMAG.Glasgow_Norms, PoS.SUBTLEX_UK)
```

The suggested matches are returned in a dataframe, filtered to be within the specified tolerances, and ordered by euclidean distance from the target word (calculated using all the numeric variables used). The closest suggested match for "elephant" is "sandwich". If we are looking for a match to fit in a sentential context, we can choose the best suitable match from this list.

```{r, echo=FALSE, class='table'}
suggested_matches %>%
  knitr::kable()
```

## Matching by Similarity

You may want to match by similarity to the target word. Thankfully this is more straightforward than in the generate pipeline (see [`control_for_map()`](advanced-stimulus-generation.html#map-functions-as-controls)).

### Orthographic similarity

Here's an example, matching "leaflet" by orthographic similarity (Levenshtein distance). We just have to calculate the similarity measure before using the `match_item()` function.

```{r, message=FALSE}
library(LexOPS)
library(stringdist)
library(dplyr)

target_word <- "interesting"

suggested_matches <- lexops %>%
  mutate(orth_sim = stringdist(string, target_word, method="lv")) %>%
  match_item(target = target_word, orth_sim = 0:3)
```

Note that some of these are misspellings or unusual words, but we could remove these by filtering (e.g. with `dplyr::filter()`) or matching (with `match_item()`) by frequency, proportion known, or familiarity ratings.

```{r, echo=FALSE, class='table'}
suggested_matches %>%
  dplyr::select(string, euclidean_distance, orth_sim) %>%
  head(10) %>%
  knitr::kable()
```

### Phonological Similarity

To match by phonological similarity, we just have to calculate the Levenshtein distance on one-letter phonemic representations, e.g. with `CMU.1letter` or `eSpeak.br_1letter`. Here we find words that are only 0 to 2 phonemic insertions, deletions, or substitutions away from "interesting".

```{r, message=FALSE}
library(LexOPS)
library(stringdist)
library(dplyr)

target_word <- "interesting"

# get the target word's pronunciation
target_word_pron <- lexops %>%
  filter(string == target_word) %>%
  pull(CMU.1letter)

# find phonologically similar words
suggested_matches <- lexops %>%
  mutate(phon_sim = stringdist(CMU.1letter, target_word_pron, method="lv")) %>%
  match_item(target_word, phon_sim = 0:2)
```

Which gives us:

```{r, echo=FALSE, class='table'}
suggested_matches %>%
  dplyr::select(string, euclidean_distance, phon_sim, CMU.1letter) %>%
  head(10) %>%
  knitr::kable()
```

